Fast Bayesian Inference for Exoplanet Discovery in Radial Velocity Data

### 2014-10-10 Paper/ms.tex
Apple pie and bayes

% Bold symbols
\renewcommand{\btheta}{\{\theta_i\}}
\newcommand{\bx}{\{x_i\}}

@@

\section{Introduction}
The number of known extrasolar planets has exploded in the last two
decades \citep{}. These discoveries have been driven by improvements in
all of the different techniques used to detect and characterise exoplanets,
including the radial velocity method \citep[e.g.][]{},
the transit method \citep[e.g.]{},
and gravitational microlensing \citep[e.g.][]{}. This paper focuses on the
radial velocity method.

Many authors have considered the problem of inferring the properties of an
exoplanetary system from observational data. With radial velocity data,
the expected signal due to an exoplanet is periodic, and the goal is to
infer the number of planets in the system, as well as their properties such
as orbital periods and eccentricities. Many
different techniques have been proposed. These techniques fall into two main
classes: i) those based on periodograms \citep[e.g.][]{}, ii) those based on
Bayesian inference, to describe the uncertainties probabilistically.

\section{Inference}

@@

The number of orbiting planets, $N$, is usually of significant interest.
Most authors have carried out the inference for $N$ by doing separate runs, one
for each possible value of $N$, and calculating the marginal likelihood
$p(\boldsymbol{D} | N)$ as a function of $N$
\citep[e.g.][]{2011MNRAS.415.2523G, 2014MNRAS.437.3540F, fengji}.

@@

The posterior distribution for the logarithms of the periods is shown in
Figure~\ref{fig:nu_oph_periods}. Because of the label-switching degeneracy,
the posterior samples for all periods were combined to make this figure, which
shows three prominent peaks. Defining the log-periods by
$S_i =  \log_{10}(T/\textnormal{1 day})$,
Figure~\ref{fig:nu_oph_periods} is a Monte Carlo representation
of the mixture distribution
\begin{eqnarray}
f(S) &=& \sum_{N=1}^{10} \frac{1}{N}\sum_{i=1}^N p(S_i | N, \boldsymbol{D}).
\end{eqnarray}

@@

\begin{thebibliography}{99}
\bibitem[\protect\citeauthoryear{Feroz 
\& Hobson}{2014}]{2014MNRAS.437.3540F} Feroz F., Hobson M.~P., 2014, MNRAS, 437, 3540

@@

\bibitem[\protect\citeauthoryear{Sato et al.}{2012}]{2012PASJ...64..135S} 
Sato B., et al., 2012, PASJ, 64, 135 
\end{thebibliography}

>>>>

% Bold symbols
\renewcommand{\btheta}{\boldsymbol{\theta}}
\newcommand{\bdata}{\boldsymbol{D}}
\newcommand{\bx}{\{x_i\}}

@@

\section{Introduction}
The number of known extrasolar planets has exploded in the last two
decades \citep{}. These discoveries have been driven by improvements in
all of the different techniques used to detect and characterise exoplanets,
including the radial velocity method \citep[e.g.][]{},
the transit method \citep[e.g.]{},
and gravitational microlensing
\citep[e.g.][]{2014ApJ...785..155B, 2014ApJ...790...14Y}. This paper focuses on the
radial velocity method.

Many authors have considered the problem of inferring the properties of an
exoplanetary system from observational data. With radial velocity data,
the expected signal due to an exoplanet is periodic, and the goal is to
infer the number of planets in the system, as well as their properties such
as orbital periods and eccentricities. Many
different techniques have been proposed. These techniques fall into two main
classes: i) those based on periodograms \citep[e.g.][]{},
and ii) those based on model fitting in the Bayesian inference framework,
to describe the uncertainties probabilistically.

\section{Inference}
Bayesian inference is the use of probability theory to describe uncertainty
\citep{}. In this framework, we approach data analysis problems by first
constructing a {\it hypothesis space}, which is the set of possible answers
to the problem we are considering. Normally, this is the set of possible
values of a vector of parameters $\btheta$ whose values we want
to know. We then assign probability distributions
called the {\it prior} and the {\it sampling distribution}. The prior
distribution $p(\btheta)$ describes
our initial uncertainty about which values of the parameters
$\btheta$ are plausible, and the
sampling distribution $p(\bdata | \btheta)$ describes
our initial uncertainty about the data set we're going to observe, as a
function of the unknown parameters $\btheta$.
When the data is known, our state of knowledge about the parameter is
updated from the prior $p(\theta)$ to the posterior distribution given by
Bayes' rule:
\begin{eqnarray}
p(\btheta | \bdata) &=&
\frac{p(\btheta)p(\bdata | \btheta)}
{p(\bdata)}
\end{eqnarray}
where the denominator, often called the {\it evidence} or
{\it marginal likelihood}, is given by:
\begin{eqnarray}
p(\bdata) = \int p(\btheta) p(\bdata | \btheta) d^N \btheta
\end{eqnarray}

@@

The number of orbiting planets, $N$, is usually of significant interest.
Most authors have carried out the inference for $N$ by doing separate runs, one
for each possible value of $N$, and calculating the marginal likelihood
$p(\bdata | N)$ as a function of $N$
\citep[e.g.][]{2011MNRAS.415.2523G, 2014MNRAS.437.3540F, fengji}.

@@

The posterior distribution for the logarithms of the periods is shown in
Figure~\ref{fig:nu_oph_periods}. Because of the label-switching degeneracy,
the posterior samples for all periods were combined to make this figure, which
shows three prominent peaks. Defining the log-periods by
$S_i =  \log_{10}(T/\textnormal{1 day})$,
Figure~\ref{fig:nu_oph_periods} is a Monte Carlo representation
of the mixture distribution
\begin{eqnarray}
f(S) &=& \sum_{N=1}^{10} \frac{1}{N}\sum_{i=1}^N p(S_i | N, \bdata).
\end{eqnarray}

@@

\begin{thebibliography}{99}

\bibitem[\protect\citeauthoryear{Bennett et 
al.}{2014}]{2014ApJ...785..155B} Bennett D.~P., et al., 2014, ApJ, 785, 155

\bibitem[\protect\citeauthoryear{Feroz 
\& Hobson}{2014}]{2014MNRAS.437.3540F} Feroz F., Hobson M.~P., 2014, MNRAS, 437, 3540

@@

\bibitem[\protect\citeauthoryear{Sato et al.}{2012}]{2012PASJ...64..135S} 
Sato B., et al., 2012, PASJ, 64, 135

\bibitem[\protect\citeauthoryear{Yee et al.}{2014}]{2014ApJ...790...14Y} 
Yee J.~C., et al., 2014, ApJ, 790, 14

\end{thebibliography}
